---
title: "Retail Analysis"
author: "DINESH RAJAN R A"
date: "6/4/2021"
output:
  html_document:
    df_print: paged
---

*Reading the files*

```{r}
customer=read.csv("C:\\Users\\DELL\\Downloads\\Customer.csv",header = TRUE,stringsAsFactors = FALSE,na.strings = c("","NA"))
products=read.csv("C:\\Users\\DELL\\Downloads\\prod_cat_info.csv",header= TRUE,stringsAsFactors = FALSE,na.strings = c("","NA"))
transaction=read.csv("C:\\Users\\DELL\\Downloads\\Transactions.csv",header = TRUE,stringsAsFactors = FALSE,na.strings = c("","NA"))
```

*Loading the libraries*

```{r}
library(dplyr)
library(ggplot2)
```

**1. Merge the datasets Customers, Product Hierarchy and Transactions as Customer_Final. Ensure to keep all customers who have done transactions with us and select the join type accordingly.**

**1. a) Merge using the base merge()**

```{r}
trans=merge(transaction,
            customer,
            by.x = "cust_id",
            by.y = "customer_Id",
            all.x = TRUE,
            all.y = FALSE)
cust=merge(trans,
           products,
           by.x=c("prod_cat_code","prod_subcat_code"),
           by.y=c("prod_cat_code","prod_sub_cat_code"),
           all.x = TRUE,
           all.y = FALSE)

cat('Dimensions of dataframe obtained through merge(): ', dim(cust))
```

*1. b) Merge using Dplyr merge functions*

```{r}
trans1=left_join(transaction,
                 customer,
                 by=c("cust_id" = "customer_Id"))
Customer_Final=left_join(trans1,
                products,
                by=c("prod_cat_code" = "prod_cat_code",
                     "prod_subcat_code" = "prod_sub_cat_code"))
cat('Dimensions of dataframe obtained through dplyr(): ', dim(Customer_Final))
```

**2. Prepare a summary report for the merged data set.**

*2. a) Get the column names and their corresponding data types.*

```{r}
str(Customer_Final)
```

*2. b) Top/Bottom 10 observations*

```{r}
head(Customer_Final,n=10)
tail(Customer_Final,n=10)
```

*2. c) “Five-number summary” for continuous variables (min, Q1, median,Q3 and max)*

```{r}
summary(select(Customer_Final,c(Qty,Rate,Tax,total_amt)))
```

*2. d) Frequency tables for all the categorical variables*

```{r}
table(select(Customer_Final,c(Gender,Store_type,prod_cat,prod_subcat)))
```


**3. Generate histograms for all continuous variables and frequency bars for categorical variables.**

*Histograms for continuous variables*

```{r}
quantity=ggplot(Customer_Final,aes(x=Qty)) + geom_histogram(bins=30,color="red",fill="orange",linetype="dashed") +
  ggtitle("HISTOGRAM FOR QUANTITY ")+
  xlab("QUANTITY")+
  ylab("COUNT")
plot(quantity)
```
```{r}
rate=ggplot(Customer_Final,aes(x=Rate)) + geom_histogram(bins=30,color="red",fill="orange",linetype="dashed") +
  ggtitle("HISTOGRAM FOR RATE ")+
  xlab("RATE")+
  ylab("COUNT")
plot(rate)
```
```{r}
tax=ggplot(Customer_Final,aes(x=Tax)) + geom_histogram(bins=30,color="red",fill="orange",linetype="dashed") +
ggtitle("HISTOGRAM FOR TAX ")+
xlab("TAX")+
ylab("COUNT")
plot(tax)
```
```{r}
total_amt=ggplot(Customer_Final,aes(x=total_amt)) + geom_histogram(bins=30,color="red",fill="orange",linetype="dashed") +
ggtitle("HISTOGRAM FOR TOTAL AMOUNT ")+
xlab("TOTAL AMOUNT")+
ylab("COUNT")
plot(total_amt)
```

*Frequency bars for continuous variables*

```{r}
quantity=ggplot(Customer_Final,aes(x=Qty)) + geom_bar(color="red",fill="orange",linetype="dashed") +
  ggtitle("FREQUENCY BAR FOR QUANTITY ")+
  xlab("QUANTITY")+
  ylab("COUNT")
plot(quantity)
```
```{r}
rate=ggplot(Customer_Final,aes(x=Rate)) + geom_bar(color="red",fill="orange",linetype="dashed") +
  ggtitle("FREQUENCY BAR FOR RATE ")+
  xlab("RATE")+
  ylab("COUNT")
plot(rate)
```
```{r}
tax=ggplot(Customer_Final,aes(x=Tax)) + geom_bar(color="red",fill="orange",linetype="dashed") +
  ggtitle("FREQUENCY BAR FOR TAX ")+
  xlab("TAX")+
  ylab("COUNT")
plot(tax)
```
```{r}
total_amt=ggplot(Customer_Final,aes(x=total_amt)) + geom_bar(color="red",fill="orange",linetype="dashed") +
  ggtitle("FREQUENCY BAR FOR TOTAL AMOUNT ")+
  xlab("TOTAL AMOUNT")+
  ylab("COUNT")
plot(total_amt)
```

**4. Calculate the following information using the merged dataset :**

*4 a) Calculate Time period of the available transaction data.*

```{r}
difftime("2014-02-28","2012-01-01",units="days")
```

*4 b) Calculate count of transactions where the total amount of transaction was negative*

```{r}
count(filter(Customer_Final,total_amt < 0))
```

**5. Analyze which product categories are more popular among females vs male customers.**

*Most popular categories among males*

```{r}
male=select(Customer_Final,c("Gender","prod_cat")) %>%
  filter(Gender=="M") %>% 
  group_by(prod_cat) %>%
  count(prod_cat)
arrange(male,desc(n))
```

*Most popular categories among Females*

```{r}
female=select(Customer_Final,c("Gender","prod_cat")) %>%
  filter(Gender=="F") %>% 
  group_by(prod_cat) %>%
  count(prod_cat)
arrange(female,desc(n))
```

*The top 3 categories for both males and females are:*

* Books

* Electronics

* Home and kitchen

**6. Which City code has the maximum customers and what was the percentage of customers from that city?**

*Percentage of customers belonging to different city codes sorted in descending order:*

```{r}
code=select(Customer_Final,'city_code')
c1=group_by(code,city_code) 
c1=count(c1)
sum=sum(c1$n)

c1=mutate(c1,percent= (n / sum) *100)
arrange(c1,desc(percent))
```

The highest percentage of customers come from city code 4 (10.51%)

**7. Which store type sells the maximum products by value and by quantity?**

*Products sold by value:*

```{r}
d1=select(Customer_Final,c('Store_type','Qty'))
d1=group_by(d1,Store_type)
d1=count(d1)
d1=arrange(d1,desc(n))
d1
```

```{r}
d2=select(Customer_Final,c('Store_type','Rate'))
d2=group_by(d2,Store_type)
d2=count(d2)
d2=arrange(d2,desc(n))
d2
```


**8. What was the total amount earned from the "Electronics" and "Clothing"categories from Flagship Stores?**

```{r}
flg=select(Customer_Final,c('Store_type','prod_cat','total_amt'))
  
flg=filter(flg,Store_type=='Flagship store')
flg=filter(flg,prod_cat==('Electronics') | prod_cat==('Clothing'))
summarise(flg,sum(total_amt))
```

**9. What was the total amount earned from "Male" customers under the "Electronics" category?**

```{r}
male_electronics=select(Customer_Final,c('Gender','prod_cat','total_amt')) %>%
  filter(Gender=="M" & prod_cat=="Electronics") %>%
  summarise(sum(total_amt))
male_electronics
```

**10. How many customers have more than 10 unique transactions, after removing all transactions which have any negative amounts?**

```{r}
uni=select(Customer_Final,c('transaction_id','cust_id','Qty','Rate','total_amt')) %>%
  filter(Qty > 0 & Rate > 0 & total_amt > 0) %>%
  count(cust_id) %>%
  filter(n >10)

count(uni)
```